---
title: "Photodiode Post-Excitation Waveform: Statistical Audit"
subtitle: "Computational reproduction, control audit, and signal-processing audit"
date: today
params:
  repo_root: ".."
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
repo <- tryCatch(params$repo_root, error = function(e) "..")
T <- file.path(repo, "tables_r", "statistical_audit")
F <- file.path(repo, "figures_r", "statistical_audit")
rd <- function(f) {
  p <- file.path(T, f)
  if (file.exists(p)) utils::read.csv(p, check.names = FALSE) else data.frame()
}
fig <- function(f) {
  p <- file.path(F, f)
  if (file.exists(p)) knitr::include_graphics(normalizePath(p)) else invisible()
}
show <- function(df, n = 30) {
  if (!nrow(df)) return(cat("_no data_"))
  knitr::kable(utils::head(df, n))
}
```

# 1. Executive summary

This audit re-implements, in R, the analysis behind the repository's claim of an
"optically induced harmonic state" in a silicon photodiode, and stress-tests the
measurement against alternative signal origins and analysis artifacts, including
60 Hz-related electrical/environmental coupling, instrument effects, spectral
leakage, peak-selection, and analysis-window choices.

**Potentially supported by the data:**

- The post-excitation capture contains a discrete **60 Hz-spaced harmonic
  ladder** (peaks at 60–600 Hz).
- The **canonical Python log-cos calculation is reproduced in R** to ~13
  significant figures (see Parity, §19).
- Particular **2:3:4 frequency ratios** occur among the detected peaks.
- The log-cos scan-max **exceeds the canonical shuffle null** (p at the Monte
  Carlo resolution floor).
- The archived runs are **condition-dependent rather than uniformly locked**:
  not every recorded run contains the same harmonic state. This is relevant to
  source attribution, but the runs are not yet a complete matched-control set.

**NOT established by the available data:**

- That optical induction uniquely caused the structure.
- That any specific environmental, conducted, magnetic, grounding, or
  instrument-coupling mechanism caused the structure. The 60 Hz spacing means
  such mechanisms remain important alternatives, but frequency coincidence
  alone does not identify the source.
- That the two exact Q = 2/3 triplets are statistically independent — both lie
  on the 60 Hz ladder and **share the 240 Hz peak**; the geometry is an
  **arithmetic consequence**, not independent statistical evidence.
- That the shuffle p-value alone is sufficient for physical source attribution;
  colored-noise and structured harmonic-ladder nulls answer different
  statistical questions.
- Any claim of **persistence** beyond the canonical 0.5 s capture, a **resonator
  quality factor**, or confirmation of WCT.

```{r}
show(rd("final_claim_matrix.csv"))
```

# 2. Experimental claim and alternatives

The scientific question is whether the post-excitation waveform contains a
reproducible condition-dependent device state distinguishable from conducted,
magnetic, environmental, grounding, and instrument coupling, as well as from
spectral leakage, peak-selection, and analysis-window effects. No single
alternative mechanism is treated as established from the present archive.

# 3. Capture inventory and prediction ledger

```{r}
show(rd("capture_registry.csv"))
```

**Data reality:** one raw canonical post-excitation `.npy`, two–three raw CSV
waveforms (post / undocumented phase), a baseline that exists **only as a peak
table**, and an incomplete set of matched raw controls.

The experimental history also included improvised electromagnetic shielding of
the photodiode and a restart/repeat in which a delayed response was observed
without entering the previously observed locked harmonic state. These are
useful condition-dependence observations, but they are not substitutes for
quantified matched controls and do not by themselves exclude conducted or
low-frequency magnetic coupling.

```{r}
fig("fig01_capture_flow.png")
show(rd("prediction_ledger_audit.csv"))
```

The prediction ledger demonstrates chronology, not correctness, and is not a
substitute for controls.

# 4. Waveform calibration and quality control

```{r}
show(rd("waveform_qc.csv"))
fig("fig02_waveform_qc.png")
```

The Siglent convention `voltage = raw_byte * vscale / 25 - offset` implies a
quantization step of `vscale/25 = 0.12 V`; the committed `.npy` values are
integer multiples of 0.12 V, consistent with an 8-bit ADC.

# 5. Baseline versus post-excitation comparison

```{r}
fig("fig03_baseline_post_waveforms.png")
fig("fig05_difference_spectrum.png")
show(rd("harmonic_effect_sizes.csv"))
```

Baseline and post amplitudes are on **different normalizations** (calibrated V
vs unnormalized FFT) and are not directly comparable; the difference spectrum
cannot be computed without the baseline raw waveform.

# 6. Spectral estimation

```{r}
show(rd("spectrum_summary.csv"))
fig("fig04_psd_comparison.png")
```

Native FFT-bin spacing is **2 Hz** (0.5 s capture). Raw FFT amplitudes are not
compared across windows without coherent-gain normalization; zero-padding is
used for interpolation only, never as true resolution.

# 7. 60 Hz-related coupling alternative

```{r}
show(rd("peak_estimates.csv"))
fig("fig07_frequency_residuals.png")
```

The detected ladder falls on integer multiples of 60 Hz, and the baseline peak
table also contains 60 Hz-coincident frequencies. This means mains-related or
environmental coupling **cannot be excluded** from the present data. It does
not establish mains pickup as the source. The archived behavior is
condition-dependent, including runs without the same locked ladder, so source
attribution requires matched raw controls rather than inference from frequency
spacing alone.

# 8. Harmonic ladder

```{r}
show(rd("harmonic_ladder.csv"))
fig("fig06_harmonic_ladder.png")
```

# 9. Q = 2/3 ratio geometry

```{r}
show(rd("ratio_geometry.csv"), 12)
show(rd("triplet_overlap.csv"))
fig("fig09_q_geometry.png")
fig("fig10_triplet_overlap.png")
```

Any integer-harmonic triplet proportional to 2:3:4 gives `Q_low = Q_high = 2/3`
by construction. The triplets (120,180,240) and (240,360,480) **share 240 Hz**
and are **not** statistically independent.

# 10. Canonical log-cos reproduction

```{r}
show(rd("logcos_results.csv"))
fig("fig11_logcos_scan.png")
fig("fig20_python_r_parity.png")
```

# 11. Stronger log-cos nulls

```{r}
show(rd("logcos_null_comparison.csv"))
fig("fig12_logcos_nulls.png")
```

The canonical shuffle null destroys frequency-local correlation. Block
permutation, stationary bootstrap, colored-noise (AR), and synthetic
harmonic-ladder nulls preserve different kinds of structure and test whether the
observed log-cos statistic is unusual under those model classes. These tests do
**not** identify the physical source of the measured waveform: statistical
compatibility with a synthetic 60 Hz-harmonic model is not equivalent to a
causal finding of mains pickup.

# 12. Window and detrending sensitivity

```{r}
show(rd("window_sensitivity.csv"), 20)
fig("fig14_window_sensitivity.png")
fig("fig13_logcos_bootstrap.png")
```

# 13. Time-frequency persistence

```{r}
show(rd("persistence_results.csv"))
fig("fig15_spectrogram.png")
fig("fig16_harmonic_persistence.png")
```

Persistence here means stability **within** the canonical 0.5 s capture only.
The broader experimental history includes subsequent condition-dependent
behavior, including a delayed response without harmonic lock, but those runs do
not yet constitute a calibrated ringdown series.

# 14. Experimental controls

```{r}
show(rd("control_comparison.csv"))
fig("fig17_control_matrix.png")
```

Missing matched control categories are listed explicitly and shown as NA. The
current archive is therefore insufficient to assign a unique physical cause to
the locked spectrum. In particular, the data do not establish either "mains
pickup" or "optically induced state" as a completed causal conclusion.

# 15. Ringdown analysis

```{r}
show(rd("ringdown_models.csv"))
```

Not established from the canonical capture: no calibrated time-resolved
post-excitation amplitude series suitable for a resonator-Q measurement.
`Q_ratio` (2/3 geometry) is distinct from `Q_resonator = pi f tau`.

# 16. Multiple-testing correction

```{r}
show(rd("multiple_testing.csv"))
```

The canonical scan-global p corrects only for the look-elsewhere effect across
k — not for capture, window, detrending, peak-threshold, ratio, or control
selection.

# 17. Null calibration

```{r}
show(rd("null_calibration.csv"))
fig("fig18_null_calibration.png")
```

# 18. Injection recovery

```{r}
show(rd("injection_recovery.csv"), 40)
fig("fig19_injection_recovery.png")
```

# 19. Python/R parity

```{r}
show(rd("python_r_parity.csv"))
```

R reproduces the deterministic canonical quantities within tolerance. Stochastic
nulls are compared by distribution, not draw-for-draw. R never calls the Python
analysis script.

# 20. Claim-status matrix

```{r}
show(rd("final_claim_matrix.csv"))
fig("fig21_evidence_dashboard.png")
```

# 21. Limitations

- Single raw canonical post-excitation capture; baseline is peak-table only.
- Incomplete matched raw control set (sham, disconnected, terminated, dark,
  channel-swap, battery, multi-device/day).
- Improvised shielding and restart/repeat observations are informative but not
  quantitatively calibrated isolation controls.
- 2 Hz FFT resolution; exact frequencies rely on sub-bin estimates.
- Stronger nulls and calibration run at development resolution unless invoked at
  full resolution.

# 22. Reproducibility information

```{r}
cat("Generated:", as.character(Sys.time()), "\n")
print(sessionInfo())
```

Exact full-resolution command:

```
Rscript R/render_photodiode_audit.R --mode waveform --null-n 10000 \
  --bootstrap-n 5000 --calibration-n 10000 --injection-n 1000 --seed 20260309
```
